### 2020-03-28 paper.tex
Add "Flowchart" Subsection

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\makeatletter
\newcommand\fs@norules{\def\@fs@cfont{\bfseries}\let\@fs@capt\floatc@ruled
    \def\@fs@pre{}%
    \def\@fs@post{}%
    \def\@fs@mid{\kern3pt}%
    \let\@fs@iftopcapt\iftrue}
\makeatother

\tikzstyle{startstop} = [ellipse,minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=white]
\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=white]
\tikzstyle{process} = [rectangle, minimum width=1cm, minimum height=1cm, text centered, draw=black, fill=white]
\tikzstyle{decision} = [diamond, minimum width=2cm, minimum height=1cm, text centered, draw=black, fill=white]
\tikzstyle{arrow} = [thick,->,>=stealth]

\begin{document}

\title{Conference Paper Title*\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
% \thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{Khaled Amgad\IEEEauthorrefmark{1},
Mahmoud Mohamad\IEEEauthorrefmark{2}, Muhanad Atef\IEEEauthorrefmark{3} and
Nader AbdAlGhani\IEEEauthorrefmark{4}}
\IEEEauthorblockA{Computer Engineering Dept.,
Faculty of Engineering, Cairo University\\
Cairo, Egypt\\
\IEEEauthorrefmark{1}khaled.mohamed98@eng-st.cu.edu.eg,
\IEEEauthorrefmark{2}mmmacmp@gmail.com,
\IEEEauthorrefmark{3}muhanad.atef23@gmail.com,
\IEEEauthorrefmark{4}nader\_abdelghani@hotmail.com}}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}

Scheduling is a crucial part of many real-time applications such as scheduling airlines and railways communications, product manufacturing processes, data packets routing in computer networks and pipe networks. Operating systems multiprocessing environments are no different than those mediums mentioned, where processes compete over CPU utilization. This brought to existence the need for scheduling algorithms to justly assign processes to available CPUs in favour of optimizing performance measures, in particular, to maximize overall CPU utilization and throughput, and to minimize response time, waiting time and turnaround time. A schedular could be preemptive where it can temporarily interrupt a process without its cooperation and assign the CPU resources to a different process with the intention to assign them back to the former process, such operation is called context switching, or it could be non-preemptive (cooperative) by not being able to context switch between processes. Various preemptive and non-preemptive scheduling disciplines exist such as, first come, first serve (FCFS), shortest job first (SJF), shortest remaining time first (SRTF), round-robin (RR), multilevel queue and multilevel feedback queue. In FCFS, processes are non-preemptively executed according to their arrival time. In SJF, the CPU is assigned to the process with the smallest burst time. SRTF is a preemptive version of SJF where at each iteration, the process with the least remaining burst time takes control of the CPU. Round-robin preemptively assigns the CPU to each process in the ready queue for a static amount of time called quantum and executes them in an FCFS fashion. Multilevel queue algorithm partitions the ready queue into several queues to which processes are perpetually assigned and are executed according to another scheduling algorithm (e.g. RR). Processes cannot move from one queue to another. However, in a multilevel feedback queue algorithm, processes that don't terminate in one queue, due to having CPU burst time more than the time quantum assigned to their particular queue, are shifted to a lower priority queue. Due to the fact that long processes eventually sink to the lowest priority queue, they are carried out using FCFS to prevent starvation. Several papers discuss different methods to optimize the algorithms mentioned, each come with their set of advantages and drawbacks. One of many is \cite{b2}, whose approach achieves better average response time but at the cost of hindering the scheduling process due to recalculating time quantum for each queue using a recurrent neural network. Our proposed approach implements the best of all worlds regarding \cite{b1} and \cite{b3} by avoiding each drawback and optimizing their advantages.

\section{Related Work}

Several papers proposed various types of approaches to improve the overall efficiency of the multilevel feedback queue scheduling algorithm. The chosen quantum time for each queue plays a major role. For instance, in \cite{b1}, the highest priority queue has been given a relatively low quantum period, and as the priority of a queue decreases, its quantum gets multiplied by a constant. Hence, it is essential to choose a proper method to compute the time quantum value to minimize response time and maximize overall performance. In \cite{b2}, an algorithm is presented for solving these drawbacks and minimizing the response time. In this algorithm, a Recurrent Neural Network (RNN) was used to determine both the number of queues and the optimized time quantum value for each queue. The RNN generates an effective model to compute the time quantum value. Their proposed intelligent version of the MLFQ offers good results, however, it suffers from a few drawbacks, the first being the direct proportionality of its network learning time and the amount of input data, and the second is the possibility of experiencing initial overhead at the first iterations of the algorithm. Our approach proposes an enhanced version of MLFQ using an optimized version of RR named shortest remaining burst round-robin (SRBRR) as in \cite{b3} which avoids the learning and overhead time in \cite{b2}. Regarding the various approaches to improve the MLFQ algorithm, those attempts dealt with starvation by assigning different quantum values to the ready queues depending on their priority. Our approach deals with starvation by boosting processes from lower priority queues to higher ones according to a specific policy.

\section{Proposed Approach}

For a multilevel feedback queue scheduling algorithm, three parameters should be considered. The first being the chosen scheduling algorithm for each queue, especially the last queue as its scheduling algorithm must help treat starvation. The second is the criteria according to which a process priority should be increased based on its waiting time in its particular ready-queue, this technique is also known as Aging. The third is the criteria according to which a process priority should be decreased. In our proposed algorithm, there are 5 queues sorted in ascending order in line with their priority number, 0 being the highest priority and 4 being the lowest. Each queue uses a modified version of the round-robin scheduling algorithm stated in \cite{b3}. In \cite{b3}, processes are sorted in ascending order according to their burst time and are assigned a time quantum that equals the median burst time of those processes. This dynamic quantum time RR algorithm provides better turnaround time and waiting time than the standard static quantum time RR algorithm whose quantum if set too short, leads to many context switches and if set too long, morphs the algorithm into an FCFS algorithm. The proposed alteration on the stated algorithm in \cite{b3} is that each queue quantum time isn't exactly the median burst time, but the burst time whose index is shifted from the median value towards the higher burst time values with a step count equal to its queue number, hence, the gradual increase of quantum time as priority decreases. For clarification, a queue with priority equal 2 has the following processes denoted by their burst time: 100, 300, 550, 600, 620, 700, 720, 900, 1200, the median value is 620, assuming we are using zero-based numbering, its index equals 4, and since we are in a queue whose priority equals 2, therefore the quantum slice value according to the proposed approach equals the burst value at index 6 (as in 4 + 2) which is 720. If the processes of a certain queue didn't terminate after assigning its queue quantum time, they are shifted downwards to the next lower priority queue. After introducing new processes into a queue, the quantum time slice is recalculated. Processes age whenever they are in a queue whose priority is one less than that currently getting scheduled and satisfies the following inequality:
\begin{equation}
1 - \frac{waiting \; time \; of \; P}{burst \; time \; of \; P} < 0.95\label{eq}
\end{equation}

Those procedures are repeated for all the generated queues until all processes reach the lowest priority queue where they are rescheduled until their completion.

\subsection{Pseudocode Flowchart of the Proposed Approach}
\vspace{5mm}
\begin{tikzpicture}[node distance=2cm]

    \node (start) [startstop] {Start};
    \node (in1) [io, below of=start, text width = 2cm] {Read Processes Data};
    \node (pro1) [process, below of=in1, text width = 3cm] {Assign processes to their respective queue};
    \node (dec1) [decision, below of=pro1, text width = 2cm, yshift=-1cm] {All queues are empty};
    \node (end) [startstop, left of=dec1, xshift=-2cm] {End};
    \node (pro2) [process, below of=dec1, text width= 3cm, yshift = -1cm] {Iterate on non-empty queues $Q$};
    \node (dec2) [decision, below of=pro2, text width = 2cm, yshift=-1cm] {A process $P_{j}$ in $Q_{i+1}$ is boostable};
    \node (pro3) [process, left of=dec2, text width= 3cm, xshift= -2.2cm] {Promote $P_{j}$ to $Q_{i}$};
    \node (pro4) [process, below of=dec2, text width= 3cm, yshift=-1.2cm] {Calculate the quantum time value $\Gamma$ of $Q_{i}$};
    \node (pro5) [process, below of=pro4, text width= 3cm] {Assign $\Gamma$ to $Q_{i}$ processes};
    \node (pro6) [process, below of=pro5, text width= 3cm] {Downgrade incomplete processes to $Q_{i+1}\,:\,i+1 < 5$};
    \node (pro7) [process, below of=pro6, text width= 3cm] {Increment waiting time of $P_{j} \notin Q_{i}$};

    \draw [arrow] (start) -- (in1);
    \draw [arrow] (in1) -- (pro1);
    \draw [arrow] (pro1) -- (dec1);
    \draw [arrow] (dec1) -- node[anchor=west, above] {Yes} (end);
    \draw [arrow] (dec1) -- node[anchor=south, left] {No} (pro2);
    \draw [arrow] (pro2) -- (dec2);
    \draw [arrow] (dec2) -- node[anchor=west,above] {Yes} (pro3);
    \draw [arrow] (pro3.north) -- ++ (0,1.3) -- ++ (4.2,0);
    \draw [arrow] (dec2) -- node[anchor=south, left] {No} (pro4);
    \draw [arrow] (pro4) -- (pro5);
    \draw [arrow] (pro5) -- (pro6);
    \draw [arrow] (pro6) -- (pro7);
    \draw [arrow] (pro7.west) -- ++ (-4.5,0) -- ++(0,16.8) -- ++(6.1,0);

    \end{tikzpicture}

\subsection{Proposed Algorithm}
\floatstyle{norules}
\restylefloat{algorithm}
\begin{algorithm}[H]
    \caption{Enhanced Multilevel FeedBack Queue Scheduling Algorithm}
    \begin{algorithmic}[1]
    \Procedure {EMLFQ}{Number of processes $n$, priority of each process $\alpha_{i}$, burst time of each process $\beta_{i}$}
    % Return  Average waiting time, throughput
    \Statex Declaration and Initialisation:
    \State Queue $Q_{i}$ where $i \in \{0, 1, 2, 3, 4\}$, waiting time $\omega_{i} = 0$ of $P_{i}$, quantum time $\Gamma$, quantum time index $\gamma$
        \For {$i = 0$ to $n$}
            \State \textbf{Assign} $P_{i}$ to $Q_{\alpha_{i}}$
        \EndFor
    \Statex Loop Process:
        \State $i = 0$
        \While {$Q_{i}$ is NOT empty}
            % Aging Processes
            \For{$j = i + 1$ to $5$}
                \If{$1-\omega_{\theta}/\beta_{\theta} < 0.95$ for $\theta \in Q_{j}$}
                    \State \textbf{Promote} $P_{\theta}$ to the current queue $Q_{i}$
                \EndIf
            \EndFor

            % Calculating the quantum time value
            \Statex Calculating the quantum time value:
            \State $ascending\_sort(\beta_{Q_{i}})$
            \State $\gamma = Q_{i}.size()/2 + i$
            \State $\Gamma = \beta_{\gamma}$ where $\beta_{\gamma} \in Q_{i}$
            % Assigning the quantum value to each process in the queue
            \Statex Assigning the Quantum Value to each Process:
            \For{$P_{j}$ in $Q_{i}$}
                \State $\beta_{j} = \beta_{j} - \Gamma$
                \If{$\beta_{j} <= 0$}
                    \State \textbf{Remove} $P_{j}$ from $Q_{i}$
                \ElsIf {$i < 4$}
                    \State \textbf{Downgrade} $P_{j}$ to $Q_{i+1}$
                \EndIf
            \EndFor
            % Increment waiting time
            \Statex Increment Waiting Time:
            \For{$j = 0$ to $n$ \textbf{and} $P_{j} \notin Q_{i}$}
                \State $\omega_{j} = \omega_{j} + 1$
            \EndFor
            \Statex Increment Queue Index
            \If{$i < 4$}
                \State $i = i + 1$
            \EndIf
        \EndWhile
     \EndProcedure
    \end{algorithmic}
\end{algorithm}

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a 
separate text file. Complete all content and organizational editing before 
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on 
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been 
formatted and styled. Do not number text heads---{\LaTeX} will do that 
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text, 
even after they have been defined in the abstract. Abbreviations such as 
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use 
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
\item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
\item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
\item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
\item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your 
equations more compact, you may use the solidus (~/~), the exp function, or 
appropriate exponents. Italicize Roman symbols for quantities and variables, 
but not Greek symbols. Use a long dash rather than a hyphen for a minus 
sign. Punctuate equations with commas or periods when they are part of a 
sentence, as in:
\begin{equation}
a+b=\gamma\label{eq}
\end{equation}

Be sure that the 
symbols in your equation have been defined before or immediately following 
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at 
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
\item The word ``data'' is plural, not singular.
\item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
\item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
\item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
\item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
\item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
\item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
\item Do not confuse ``imply'' and ``infer''.
\item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
\item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
\item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A 
minimum of one author is required for all articles. Author names 
should be listed starting from left to right and then moving down to the 
next line. This is the author sequence that will be used in future citations 
and by indexing services. Names should not be listed in columns nor group by 
affiliation. Please keep your affiliations as succinct as possible (for 
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through 
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not 
topically subordinate to each other. Examples include Acknowledgments and 
References and, for these, the correct style to use is ``Heading 5''. Use 
``figure caption'' for your Figure captions, and ``table head'' for your 
table title. Run-in heads, such as ``Abstract'', will require you to apply a 
style (in this case, italic) in addition to the style provided by the drop 
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For 
example, the paper title is the primary text head because all subsequent 
material relates and elaborates on this one topic. If there are two or more 
sub-topics, the next level head (uppercase Roman numerals) should be used 
and, conversely, if there are not at least two sub-topics, then no subheads 
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and 
bottom of columns. Avoid placing them in the middle of columns. Large 
figures and tables may span across both columns. Figure captions should be 
below the figures; table heads should appear above the tables. Insert 
figures and tables after they are cited in the text. Use the abbreviation 
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
\caption{Table Type Styles}
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
\cline{2-4} 
\textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
\hline
copy& More table copy$^{\mathrm{a}}$& &  \\
\hline
\multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
\end{tabular}
\label{tab1}
\end{center}
\end{table}

\begin{figure}[htbp]
\centerline{\includegraphics{Assets/fig1.png}}
\caption{Example of a figure caption.}
\label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words 
rather than symbols or abbreviations when writing Figure axis labels to 
avoid confusing the reader. As an example, write the quantity 
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including 
units in the label, present them within parentheses. Do not label axes only 
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization 
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of 
quantities and units. For example, write ``Temperature (K)'', not 
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without 
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B. 
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor 
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The 
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at 
the bottom of the column in which it was cited. Do not put footnotes in the 
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use 
``et al.''. Papers that have not been published, even if they have been 
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers 
that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
Capitalize only the first word in a paper title, except for proper nouns and 
element symbols.

For papers published in translation journals, please give the English 
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
\bibitem{b1} H.S. Behera, Reena Kumari Naik, Suchilagna Parida, ``Improved multilevel feedback queue scheduling using dynamic time quantum and its performance analysis'', \textit{International Journal of Computer Science and Information Technologies}, vol. 3, no. 2, 2012, pp. 3801-3807.
\bibitem{b2} MohammadReza EffatParvar, Karim Faez, Mehdi EffatParvar, Mehdi Zarei, Saeed Safari, ``An intelligent MLFQ scheduling algorithm (IMLFQ) with fault tolerant mechanism'', \textit{Sixth International Conference on Intelligent Systems Design and Applications}, vol. 3, 2006, pp. 80–85.
\bibitem{b3} Rakesh Mohanty, H. S. Behera, Khusbu Patwari, Monisha Dash, ``Design and performance evaluation of a new proposed shortest remaining burst round robin (SRBRR) scheduling algorithm'', \textit{International Symposium on Computer Engineering \& Technology}, vol. 17, 2010.
%\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
% \bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE templates contain guidance text for composing and formatting papers. Please ensure that all template text is removed from your paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}